---
title: "Arsenic - 02: Taxanomy Analysis"
author: "Jan Waelchli"
geometry: margin=2cm
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE, echo=F, warning=F, message=F,}

##clear the object from memory
rm(list=ls())

#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")
options(tinytex.verbose = TRUE)

#set seed
set.seed(100)

## set source to file location
library(rstudioapi)
setwd(dirname(getActiveDocumentContext()$path))

## load functions
source("functions/functions.R")
functions(path="functions/")

## installs (if necessary) and loads libraries
libraries()

```

```{r import, echo=F, message=F, warning=F}

# Import

# set wd to RDS files
setwd("interim/01_Import_Normalization")

#Import RDS files
bDESIGN <- readRDS("bDESIGN.RDS")
bTAX <- readRDS("bTAX.RDS")
bDAT_norm <- readRDS("bDAT_norm.RDS")
bTAX_level_cols_phyla <- readRDS("bTAX_level_cols_phyla.RDS")

fDESIGN <- readRDS("fDESIGN.RDS")
fTAX <- readRDS("fTAX.RDS")
fDAT_norm <- readRDS("fDAT_norm.RDS")
fTAX_level_cols_phyla <- readRDS("fTAX_level_cols_phyla.RDS")

```

```{r bac phyla, echo=F, warning=F, message=F}

# Data as PHYLOSEQ object

#change order for melting later
bTAX <- bTAX[,c("labels", "kingdom", "phylum", "class", "order", "family", "genus", "ASV_ID", "cols_phyla")]

#create phyloseq::phyloseq objects
bDESIGN <- as.data.frame(bDESIGN)
rownames(bDESIGN) <- bDESIGN$bac_FLD
bDAT_norm <- as.matrix(bDAT_norm)
bTAX <- as.matrix(bTAX[rownames(bDAT_norm),])
bPHYSEQ <- phyloseq::phyloseq(sample_data(bDESIGN),
                   otu_table(bDAT_norm, taxa_are_rows=T),
                   tax_table(bTAX))

# agglomerate data by labels 
bPHYSEQ_phyla <- phyloseq::taxglom(bPHYSEQ, "labels") # merge by 'labels' 
bPHYSEQ_phyla <- phyloseq::transform_sample_counts(bPHYSEQ_phyla, function(x) 100 * x/sum(x)) 

# melt phyloseq::phyloseq object
bPHYSEQ_phyla_melt <- phyloseq::psmelt(bPHYSEQ_phyla) 

# Defining ASV colors by labels
bPHYSEQ_phyla_melt$labels <- as.factor(bPHYSEQ_phyla_melt$labels)
bPHYSEQ_phyla_melt$cols <- as.character(bPHYSEQ_phyla_melt$labels)

# attributing previously assigned colors
# TAX_level_cols_phyla
for(i in names(bTAX_level_cols_phyla)[names(bTAX_level_cols_phyla) %in% levels(bPHYSEQ_phyla_melt$labels)]){
  bPHYSEQ_phyla_melt[bPHYSEQ_phyla_melt$labels==paste(i), ]$cols <- bTAX_level_cols_phyla[paste(i)]
}

# Defining high abundant Phyla
# Phyla with MEAN abundances higher than 1% relative abundances
bPHYSEQ_phyla_abu <-  rownames(otu_table(bPHYSEQ_phyla))[apply(otu_table(bPHYSEQ_phyla), 1, mean, na.rm=T) > 1]
bPHYSEQ_phyla_abuP <- tax_table(bPHYSEQ_phyla)[rownames(tax_table(bPHYSEQ_phyla)) %in% bPHYSEQ_phyla_abu, "labels"]

# Defining low abundant Phyla
# Phyla with MEAN abundances lower than 1% relative abundances
bPHYSEQ_phyla_low <-  rownames(otu_table(bPHYSEQ_phyla))[apply(otu_table(bPHYSEQ_phyla), 1, mean, na.rm=T) < 1]
bPHYSEQ_phyla_lowP <- tax_table(bPHYSEQ_phyla)[rownames(tax_table(bPHYSEQ_phyla)) %in% bPHYSEQ_phyla_low, "labels"]

# subsetting the color vector to abundant labels and classes
# delete labels name of low-abundant phyla (for plot) and put them at the bottom
bPHYSEQ_phyla_melt$labels_2 <- as.character(bPHYSEQ_phyla_melt$labels)
bPHYSEQ_phyla_melt[bPHYSEQ_phyla_melt$labels_2 %in% bPHYSEQ_phyla_lowP, ]$labels_2 <- "Low abundant phyla"
bPHYSEQ_phyla_melt$labels_2 <- as.factor(bPHYSEQ_phyla_melt$labels_2)

# color matrix for plot
bcol_class <- bPHYSEQ_phyla_melt$cols
names(bcol_class) <- bPHYSEQ_phyla_melt$labels_2
bcol_class[names(bcol_class)=="Low abundant phyla"] <- "lightgrey"

```

```{r fun phyla, echo=F, warning=F, message=F}

# Data as PHYLOSEQ object

#change order for melting later
fTAX <- fTAX[,c("labels", "kingdom", "phylum", "class", "order", "family", "genus", "species", "ASV_ID", "cols_phyla")]

#create phyloseq::phyloseq objects
fDESIGN <- as.data.frame(fDESIGN)
rownames(fDESIGN) <- fDESIGN$fun_FLD
fDAT_norm <- as.matrix(fDAT_norm)
fTAX <- as.matrix(fTAX[rownames(fDAT_norm),])
fPHYSEQ <- phyloseq::phyloseq(sample_data(fDESIGN),
                   otu_table(fDAT_norm, taxa_are_rows=T),
                   tax_table(fTAX))

# agglomerate data by labels 
fPHYSEQ_phyla <- phyloseq::taxglom(fPHYSEQ, "labels") # merge by 'labels' 
fPHYSEQ_phyla <- phyloseq::transform_sample_counts(fPHYSEQ_phyla, function(x) 100 * x/sum(x)) 

# melt phyloseq::phyloseq object
fPHYSEQ_phyla_melt <- phyloseq::psmelt(fPHYSEQ_phyla) 

# Defining ASV colors by labels
fPHYSEQ_phyla_melt$labels <- as.factor(fPHYSEQ_phyla_melt$labels)
fPHYSEQ_phyla_melt$cols <- as.character(fPHYSEQ_phyla_melt$labels)

# attributing previously assigned colors
# TAX_level_cols_phyla
for(i in names(fTAX_level_cols_phyla)[names(fTAX_level_cols_phyla) %in% levels(fPHYSEQ_phyla_melt$labels)]){
  fPHYSEQ_phyla_melt[fPHYSEQ_phyla_melt$labels==paste(i), ]$cols <- fTAX_level_cols_phyla[paste(i)]
}

# Defining high abundant Phyla
# Phyla with MEAN abundances higher than 1% relative abundances
fPHYSEQ_phyla_abu <-  rownames(otu_table(fPHYSEQ_phyla))[apply(otu_table(fPHYSEQ_phyla), 1, mean, na.rm=T) > 1]
fPHYSEQ_phyla_abuP <- tax_table(fPHYSEQ_phyla)[rownames(tax_table(fPHYSEQ_phyla)) %in% fPHYSEQ_phyla_abu, "labels"]

# Defining low abundant Phyla
# Phyla with MEAN abundances lower than 1% relative abundances
fPHYSEQ_phyla_low <-  rownames(otu_table(fPHYSEQ_phyla))[apply(otu_table(fPHYSEQ_phyla), 1, mean, na.rm=T) < 1]
fPHYSEQ_phyla_lowP <- tax_table(fPHYSEQ_phyla)[rownames(tax_table(fPHYSEQ_phyla)) %in% fPHYSEQ_phyla_low, "labels"]

# subsetting the color vector to abundant labels and classes
# delete labels name of low-abundant phyla (for plot) and put them at the bottom
fPHYSEQ_phyla_melt$labels_2 <- as.character(fPHYSEQ_phyla_melt$labels)
fPHYSEQ_phyla_melt[fPHYSEQ_phyla_melt$labels_2 %in% fPHYSEQ_phyla_lowP, ]$labels_2 <- "Low abundant phyla"
fPHYSEQ_phyla_melt$labels_2 <- as.factor(fPHYSEQ_phyla_melt$labels_2)

# color matrix for plot
fcol_class <- fPHYSEQ_phyla_melt$cols
names(fcol_class) <- fPHYSEQ_phyla_melt$labels_2

```

# Taxonomy

## Phyla abundance plot

We get an overview over the abundance of bacterial taxonomy by showing the most abundant phyla for each sample.

### Figure S14.1 | Bacteria: Phylum level taxonomy

```{r bac fig, echo=F, warning=F, message=F, fig.height=13, fig.width=15}

#sort
bPHYSEQ_phyla_melt <- bPHYSEQ_phyla_melt[order(bPHYSEQ_phyla_melt$genotype),]
bPHYSEQ_phyla_melt <- bPHYSEQ_phyla_melt[order(bPHYSEQ_phyla_melt$arsenic),]

#factor
#unique has another order than levels()
bPHYSEQ_phyla_melt$labels_2 <- factor(bPHYSEQ_phyla_melt$labels_2, levels = unique(bPHYSEQ_phyla_melt$labels_2)[1:14])#custom reordering
bPHYSEQ_phyla_melt$sample_ID <- factor(bPHYSEQ_phyla_melt$sample_ID, levels = unique(bPHYSEQ_phyla_melt$sample_ID))#factor in the order as the file is ordered now


#plot
fig3.1 <- ggplot(bPHYSEQ_phyla_melt, aes_string(x="sample_ID", y="Abundance", fill="labels_2")) + 
  geom_bar(stat="identity")+
  xlab("") + 
  ylab("Relative abundance [%]") +
  scale_colour_manual(values=bcol_class)+
  scale_fill_manual(values=bcol_class) +
  guides(fill=guide_legend(title="Phylum"))+
  facet_grid( ~ arsenic, scale="free_x", space="free_x") +
  geom_text(data=bPHYSEQ_phyla_melt, aes(label=genotype, y=-10), angle=90, check_overlap=T, size=3, hjust = 0.25) + 
  theme_bw()+
  ggtitle("Bacteria: Taxonomy")+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

#png("figs_for_paper/fig3.1.png",4200,3000, res=600)
print(fig3.1)
#dev.off()

```

\vspace{5mm}

### Figure S14.2 | Fungi: Phylum level taxonomy

```{r fun fig, echo=F, warning=F, message=F, fig.height=13, fig.width=15}

#sort
fPHYSEQ_phyla_melt <- fPHYSEQ_phyla_melt[order(fPHYSEQ_phyla_melt$genotype),]
fPHYSEQ_phyla_melt <- fPHYSEQ_phyla_melt[order(fPHYSEQ_phyla_melt$arsenic),]

fcol_class[names(fcol_class)=="Low abundant phyla"] <- "lightgrey"

#factor
fPHYSEQ_phyla_melt$labels_2 <- factor(fPHYSEQ_phyla_melt$labels_2, levels = unique(fPHYSEQ_phyla_melt$labels_2)[c(1:9,11:12,10)])#custom reordering
fPHYSEQ_phyla_melt$sample_ID <- factor(fPHYSEQ_phyla_melt$sample_ID, levels = unique(fPHYSEQ_phyla_melt$sample_ID))#factor in the order as the file is ordered now

#plot
fig3.2 <- ggplot(fPHYSEQ_phyla_melt, aes_string(x="sample_ID", y="Abundance", fill="labels_2")) + 
  geom_bar(stat="identity")+
  xlab("") + 
  ylab("Relative abundance [%]") +
  scale_colour_manual(values=fcol_class)+
  scale_fill_manual(values=fcol_class) +
  guides(fill=guide_legend(title="Phylum"))+
  facet_grid( ~ arsenic, scale="free_x", space="free_x") +
  geom_text(data=fPHYSEQ_phyla_melt, aes(label=genotype, y=-10), angle=90, check_overlap=T, size=3, hjust = 0.25) + 
  theme_bw()+
  ggtitle("Fungi: Taxonomy")+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

#png("figs_for_paper/fig3.2.png",4200,3000, res=600)
print(fig3.2)
#dev.off()

```

\pagebreak

## Effect of all factors on phyla abundances

We test if there are any difference between the phyla abundances between genotypes, arsenic-treatments or their interaction by performing a PERMNOVA (permutations = 999).

### Bacteria

```{r bac PERMANOVA, echo=F, warning=F, message=F}

### BACTERIA

#get phyla abu per sample
bphyla_sum <- aggregate(x=bPHYSEQ_phyla_melt$Abundance, by=list(bPHYSEQ_phyla_melt$labels_2, bPHYSEQ_phyla_melt$Sample), FUN=sum) %>%
  tidyr::pivot_wider(names_from = Group.2, values_from = x, ) %>%
  as.data.frame()
rownames(bphyla_sum) <- bphyla_sum$Group.1
bphyla_sum <- bphyla_sum[,colnames(bphyla_sum) != "Group.1"]

#should be 100% per sample
#colSums(bphyla_sum)

#PERMANOVA

bdist <- vegan::vegdist(t(bphyla_sum), method="bray")
bdist_paov <- vegan::adonis2(bdist ~ arsenic*genotype, data=bDESIGN)
pander(bdist_paov, caption="Bacteria: PERMANOVA")

```

### Fungi

```{r fun PERMANOVA, echo=F, warning=F, message=F}

### FUNGI

#get phyla abu per sample
fphyla_sum <- aggregate(x=fPHYSEQ_phyla_melt$Abundance, by=list(fPHYSEQ_phyla_melt$labels_2, fPHYSEQ_phyla_melt$Sample), FUN=sum) %>%
  tidyr::pivot_wider(names_from = Group.2, values_from = x, ) %>%
  as.data.frame()
rownames(fphyla_sum) <- fphyla_sum$Group.1
fphyla_sum <- fphyla_sum[,colnames(fphyla_sum) != "Group.1"]

#should be 100% per sample
#colSums(fphyla_sum)

#PERMANOVA

fdist <- vegan::vegdist(t(fphyla_sum), method="bray")
fdist_paov <- vegan::adonis2(fdist ~ arsenic*genotype, data=fDESIGN)
pander(fdist_paov, caption="Fungi: PERMANOVA")

#export for paper

#combine tables
df1 <- cbind(Taxa="Bacteria", Factor=rownames(bdist_paov), bdist_paov)
df2 <- cbind(Taxa="Fungi", Factor=rownames(fdist_paov), fdist_paov)
df_all <- rbind(df1, df2)
rownames(df_all) <- NULL
rm(df1, df2)

#save
#write.csv(df_all, "figs_for_paper/phyla_abundance.csv", row.names = F)


```

**Conclusion:** No differences found.

\pagebreak

```{r export RDA files, echo=F, message=F, warning=F}

# create directory
dir.create("interim")
dir.create("interim/02_Taxa_Analysis")

## set output directory
setwd("interim/02_Taxa_Analysis")

#save objects needed in the following scripts as RDA
saveRDS(bPHYSEQ, "bPHYSEQ.RDS")
saveRDS(bPHYSEQ_phyla_melt, "bPHYSEQ_phyla_melt.RDS")
saveRDS(fPHYSEQ, "fPHYSEQ.RDS")
saveRDS(fPHYSEQ_phyla_melt, "fPHYSEQ_phyla_melt.RDS")

```

