---
title: "Arsenic - 03: Alpha Diversity"
author: "Jan Waelchli"
geometry: margin=2cm
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE, echo=F, warning=F, message=F,}

##clear the object from memory
rm(list=ls())

#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")
options(tinytex.verbose = TRUE)

#set seed
set.seed(100)

## set source to file location
library(rstudioapi)
setwd(dirname(getActiveDocumentContext()$path))

## load functions
source("functions/functions.R")
functions(path="functions/")

## installs (if necessary) and loads libraries
libraries()

```

```{r import, echo=F, message=F, warning=F}

# Import

# set wd to RDS files
setwd("interim/01_Import_Normalization")

#cImport RDS files
bDESIGN <- readRDS("bDESIGN.RDS")
bDAT <- readRDS("bDAT.RDS")
bDAT_norm <- readRDS("bDAT_norm.RDS")

fDESIGN <- readRDS("fDESIGN.RDS")
fDAT <- readRDS("fDAT.RDS")
fDAT_norm <- readRDS("fDAT_norm.RDS")

level_cols_genotype <- readRDS("level_cols_genotype.RDS")
level_cols_genotype <- readRDS("level_cols_genotype.RDS")

```

# Alpha diversity

We answer the following questions for the alpha diversity in each substrate:

-   **Q1: Has arsenic changed the beta diversity?**
-   **Q2: Is beta diversity different between the genotypes?**
-   **Q3: Are there differences in beta diversity between the different genotypes after treating plants with arsenic?**

```{r a-div shannon, warning=F, echo=F, message=F, eval=T}

#Bacteria
asv_diversity <- vegan::diversity(bDAT_norm, index="shannon", MARGIN=2)
alpha_diversity_bDAT_summary <- cbind(asv_diversity, bDESIGN)
rm(asv_diversity)

#Fungi
asv_diversity <- vegan::diversity(fDAT_norm, index="shannon", MARGIN=2)
alpha_diversity_fDAT_summary <- cbind(asv_diversity, fDESIGN)
rm(asv_diversity)

```

## Method 

We calculate the Shannon diversity for each sample with the normalized data. 

## Genotype*Arsenic Effect

We investigate the effect on alpha diversity by the factors of genotype, arsenic and the interaction between them. We model the alpha diversity against these factors in an aov-model and perform a F-Test.

```{r a-div overview, warning=F, echo=F, message=F, eval=T}

#BACTERIA
balpha_aov <- aov(asv_diversity ~ (genotype + arsenic)^2, data=alpha_diversity_bDAT_summary) #^2 two way interactions
pander(summary(balpha_aov), caption = "Bacteria: F test")

#FUNGI
falpha_aov <- aov(asv_diversity ~ (genotype + arsenic)^2, data=alpha_diversity_fDAT_summary)
pander(summary(falpha_aov), caption = "Fungi: F test")

#export for paper

#combine tables
df1 <- cbind(Taxa="Bacteria", Factor=rownames(summary(balpha_aov)[[1]]), summary(balpha_aov)[[1]])
df2 <- cbind(Taxa="Fungi", Factor=rownames(summary(falpha_aov)[[1]]), summary(falpha_aov)[[1]])
df_all <- rbind(df1, df2)
rownames(df_all) <- NULL
rm(df1, df2)

#save
#write.csv(df_all, "figs_for_paper/alpha_diversity.csv", row.names = F)

```

```{r a-div Arsenic, warning=F, echo=F, message=F, eval=T}

#BACTERIA
balpha_aov_tukey <- emmeans::emmeans(balpha_aov, c("genotype", "arsenic"))
balpha_aov_tukey_letters <- multcomp::cld(balpha_aov_tukey, Letter="abcdefghi", alpha=0.05)
# means
balpha_means <- aggregate(alpha_diversity_bDAT_summary$asv_diversity,list(alpha_diversity_bDAT_summary$genotype, alpha_diversity_bDAT_summary$arsenic), mean)
# table
balpha_aov_tukey_letters <-balpha_aov_tukey_letters[order(with(balpha_aov_tukey_letters, genotype)),]
balpha_aov_tukey_letters <-balpha_aov_tukey_letters[order(with(balpha_aov_tukey_letters, arsenic)),]
balpha_means <- balpha_means[order(with(balpha_means, Group.1, Group.2)),]
balpha_table <- data.frame(genotype=balpha_means[,1], arsenic=balpha_means[,2], Shanon=balpha_means[,3], letters=balpha_aov_tukey_letters[,8])
#pander(balpha_table, caption = "Bacteria: alpha diversity")

#FUNGI
falpha_aov_tukey <- emmeans::emmeans(falpha_aov, c("genotype", "arsenic"))
falpha_aov_tukey_letters <- multcomp::cld(falpha_aov_tukey, Letter="abcdefghi", alpha=0.05)
# means
falpha_means <- aggregate(alpha_diversity_fDAT_summary$asv_diversity,list(alpha_diversity_fDAT_summary$genotype, alpha_diversity_fDAT_summary$arsenic), mean)
# table
falpha_aov_tukey_letters <-falpha_aov_tukey_letters[order(with(falpha_aov_tukey_letters, genotype)),]
falpha_aov_tukey_letters <-falpha_aov_tukey_letters[order(with(falpha_aov_tukey_letters, arsenic)),]
falpha_means <- falpha_means[order(with(falpha_means, Group.1, Group.2)),]
falpha_table <- data.frame(genotype=falpha_means[,1], arsenic=falpha_means[,2], Shanon=falpha_means[,3], letters=falpha_aov_tukey_letters[,8])
#pander(falpha_table, caption = "Fungi: alpha diversity")

```

### Figure 5 \| Genotype*Arsenic effect on alpha diversity

\vspace{5mm}

```{r plot a-div Plastic, echo=F, message=F, warning=F, eval=T}

#combine data.frames
alpha_diversity_bDAT_summary$taxa <- "Bacteria"
alpha_diversity_fDAT_summary$taxa <- "Fungi"
alpha_diversity_DAT_summary <- rbind(alpha_diversity_bDAT_summary, alpha_diversity_fDAT_summary)

#data frame for letters
alpha_table <- rbind(balpha_table, falpha_table)
alpha_table$taxa <- rep(c("Bacteria", "Fungi"), each=4)
alpha_table <- alpha_table[order(alpha_table$taxa, alpha_table$arsenic),] #make sure to have the same order as plot!
alpha_table$taxa <- factor(alpha_table$taxa, levels=c("Bacteria", "Fungi"))
alpha_table$letters <- trimws(alpha_table$letters) #remove whitespaces
alpha_table$x <- rep(c(0.8,1.2,1.8,2.2), 2)
alpha_table$y <- c(
tapply(alpha_diversity_bDAT_summary$asv_diversity, list(alpha_diversity_bDAT_summary$genotype, alpha_diversity_bDAT_summary$arsenic), max)+0.1,
tapply(alpha_diversity_fDAT_summary$asv_diversity, list(alpha_diversity_fDAT_summary$genotype, alpha_diversity_fDAT_summary$arsenic), max)+0.1)

#plot
plot <- ggplot(alpha_diversity_DAT_summary, aes(y=asv_diversity, x=arsenic, fill=genotype)) +
            geom_boxplot(position=position_dodge2(width=0.75, preserve="single"),outlier.colour = NA) +
            geom_jitter(size=1, position=position_jitterdodge(jitter.width=0.3,dodge.width = 0.75), alpha=0.3, aes(col=genotype))+
            theme_bw() +
            theme(legend.position="none") +
            ylab("Shanon Diversity")+
            xlab("")+
            scale_fill_manual(values=level_cols_genotype) +
            scale_color_manual(values=c("black","dimgrey","grey")) +
            facet_grid(~ taxa, scales = "free", space = "free")+
            theme(axis.text.x = element_text(angle = 0, vjust = 0.5, size=7), strip.text = element_text(size = 6)) +
            labs(title = "Alpha diversity", subtitle = "alpha diversity ~ genotype * arsenic") +
            geom_text(data = alpha_table, aes(x=x, y=y, label=letters),size=4, col="dimgrey")



#legend position
fig5 <- cowplot::plot_grid(plot + theme(legend.position="bottom"))

#print
#png("figures/fig5.png",6400, 4000, res = 600)
fig5
#dev.off()

```

\vspace{5mm}

**Conclusion:** No effect has been found.

\pagebreak

```{r export RDA files, echo=F, warning=F, message=F}

# Export

# create directory
dir.create("interim")
dir.create("interim/03_Alpha_Diversity")

## set output directory
setwd("interim/03_Alpha_Diversity")

#save objects needed in the following scripts
saveRDS(alpha_diversity_bDAT_summary, "alpha_diversity_bDAT_summary.RDS")
saveRDS(alpha_diversity_fDAT_summary, "alpha_diversity_fDAT_summary.RDS")

```
