---
title: "Arsenic - 04: Beta Diversity"
author: "Jan Waelchli"
geometry: margin=2cm
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE, echo=F, warning=F, message=F,}

##clear the object from memory
rm(list=ls())

#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")
options(tinytex.verbose = TRUE)

#set seed
set.seed(100)

## set source to file location
library(rstudioapi)
setwd(dirname(getActiveDocumentContext()$path))

## load functions
source("functions/functions.R")
functions(path="functions/")

## installs (if necessary) and loads libraries
libraries()

```

```{r paths, echo=F, message=F, warning=F}

# Import

#set wd to RDS files
setwd("interim/01_Import_Normalization")

#Import RDS files
bDESIGN <- readRDS("bDESIGN.RDS")
bDAT_norm <- readRDS("bDAT_norm.RDS")

fDESIGN <- readRDS("fDESIGN.RDS")
fDAT_norm <- readRDS("fDAT_norm.RDS")

level_cols_genotype <- readRDS("level_cols_genotype.RDS")
level_cols_arsenic <- readRDS("level_cols_arsenic.RDS")

bPHYSEQ <- readRDS("../02_Taxa_Analysis/bPHYSEQ.RDS")
fPHYSEQ <- readRDS("../02_Taxa_Analysis/fPHYSEQ.RDS")

#easier to write, from here on we only use the rarefied datasets
bDAT <- bDAT_norm
fDAT <- fDAT_norm

```

# Beta diversity

We answered the following question for the bacterial and fungal beta diversity in each compartment:

-   **Q1: Has arsenic changed the beta diversity?**
-   **Q2: Is beta diversity different between the genotypes?**
-   **Q3: Are there differences in beta diversity between the different genotypes after treating plants with arsenic?**


## Method

First we use the function 'adonis()' (package vegan) to analyze the beta diversity with a PERMANOVA (permutations = 999). Then, we graphically represent the beta diversity with a PCoA (unconstrained ordination) and a CAP plot (constrained ordination).

\vspace{5mm}

## Genotype*Arsenic Effect

We investigate the full model to see which factors alters the beta diversity.

```{r PERMANOVA all, warning=F, echo=F, eval=T}

#BACTERIA
bdist <- vegan::vegdist(t(bDAT), method="bray")
bdist_paov <- vegan::adonis2(bdist ~ (genotype + arsenic)^2, data=bDESIGN)
pander(bdist_paov, caption="Bacteria: PERMANOVA")

#FUNGI
fdist <- vegan::vegdist(t(fDAT), method="bray")
fdist_paov <- vegan::adonis2(fdist ~ (genotype + arsenic)^2, data=fDESIGN)
pander(fdist_paov, caption="Fungi: PERMANOVA")

```

\pagebreak

### Figure 6.1 \| PCoA - genotype:arsenic effect on beta diversity

\vspace{5mm}

```{r functions PCoA, echo=F, warning=F, eval=T}

#function to get centroid lines
centroid_lines <- function(PHYSEQ, dist_matrix, group, axes=c(1,2)){
  #get ordination values
  lines <- as.data.frame(phyloseq::plot_ordination(PHYSEQ, dist_matrix, axes=axes, justDF=T))
  colnames(lines)[1:2] <- c("Axis.1","Axis.2")
  lines <- lines[,colnames(lines) %in% c("Axis.1","Axis.2",group)]#subset
  #get centroids (means of each group)
  means1 <- aggregate(lines$Axis.1, by=as.list(lines[,colnames(lines) %in% group, drop=F]), FUN=mean)
  means2 <- aggregate(lines$Axis.2, by=as.list(lines[,colnames(lines) %in% group, drop=F]), FUN=mean)
  #concat groups (for multiple groups)
  lines$concact <- lines[,colnames(lines) %in% group,drop=F] %>% apply(2, as.character) %>% apply(1, paste, collapse="")
  means1$concact <- means1[,colnames(means1) %in% group,drop=F] %>% apply(2, as.character) %>% apply(1, paste, collapse="")
  means2$concact <- means2[,colnames(means2) %in% group,drop=F] %>% apply(2, as.character) %>% apply(1, paste, collapse="")
  #add means to data.frame
  lines$Axis.1.means <- NA
  lines$Axis.2.means <- NA
  group_values <- unique(lines$concact)
  for (i in group_values) {
    lines$Axis.1.means[lines$concact == i] <- means1$x[means1$concact == i]
    lines$Axis.2.means[lines$concact == i] <- means2$x[means1$concact == i]
  }
  #subset df
  lines <- lines[,colnames(lines) %in% c("Axis.1","Axis.2","Axis.1.means","Axis.2.means")]
  #return
  return(lines)
}

#scree plot
#scree plot of the first 10 MDS
scree_plot <- function(ordination, subset=1:10){
  scree <- plot_scree(ordination)
  df <- data.frame(x=scree$data$axis, y=scree$data$eigenvalue,label=paste0(round(scree$data$eigenvalue*100,2), "%"))
  df$x <- factor(df$x, levels=unique(df$x))
  scree <- ggplot(df[subset,], aes(x=x, y=y)) + 
                geom_bar(stat="identity") + 
                ylab("eigenvalue")+
                xlab("component number")+
                ylim(0,max(df$y)+0.1)+
                theme_bw()+
                geom_text(aes(label = label),vjust=-0.2, size=2)+
                theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=7))
  #return
  return(scree)
}

```

```{r genotype arsenic PCoA, echo=F, warning=F, eval=T, fig.height=9, fig.width=12}

#BACTERIA
bPCoA_bray <- phyloseq::ordinate(bPHYSEQ, method="PCoA", distance="bray")
p0 <- phyloseq::plot_ordination(bPHYSEQ, bPCoA_bray, color= "genotype", shape="arsenic", axes = c(2,3))
blines <- centroid_lines(PHYSEQ = bPHYSEQ, dist_matrix = bPCoA_bray, group = c("genotype","arsenic"), axes = c(2,3))
p0 <- p0 + geom_segment(aes(xend=blines$Axis.1, yend=blines$Axis.2), x = blines$Axis.1.means, y = blines$Axis.2.means, alpha=0.3, lwd=1)
p0 <- p0 + geom_point(size=3)
#p0 <- p0 + geom_text(aes(label=bPHYSEQ@sam_data$bac_FLD))
p0 <- p0 + scale_color_manual(values=level_cols_genotype)
p0 <- p0 + theme_bw()
p0 <- p0 + ggtitle("Bacteria - PCoA: Genotype:Arsenic effect")
fig_bac <- p0


#FUNGI
fPCoA_bray <- phyloseq::ordinate(fPHYSEQ, method="PCoA", distance="bray")
p0 <- phyloseq::plot_ordination(fPHYSEQ, fPCoA_bray, color= "genotype", shape="arsenic", axes = c(2,3))
flines <- centroid_lines(PHYSEQ = fPHYSEQ, dist_matrix = fPCoA_bray, group = c("genotype","arsenic"), axes = c(2,3))
p0 <- p0 + geom_segment(aes(xend=flines$Axis.1, yend=flines$Axis.2), x = flines$Axis.1.means, y = flines$Axis.2.means, alpha=0.3, lwd=1)
p0 <- p0 + geom_point(size=3)
#p0 <- p0 + geom_text(aes(label=fPHYSEQ@sam_data$fun_FLD))
p0 <- p0 + scale_color_manual(values=level_cols_genotype)
p0 <- p0 + theme_bw()
p0 <- p0 + ggtitle("Fungi - PCoA: Genotype:Arsenic effect")
fig_fun <- p0

#combine the two plots (library cowplot)
fig_PCoA <- cowplot::plot_grid(fig_bac + theme(legend.position="none"),
               fig_fun + theme(legend.position="none") + labs(colour=""),
               #labels=c("Bacteria","Fungi"),
               align = "hv",
               nrow=1)

#add legend
legend <- cowplot::get_legend(p0 + theme(legend.position="bottom") + labs(fill="Plastic"))
fig_PCoA <- cowplot::plot_grid(fig_PCoA, legend, nrow=2, rel_heights=c(1, .1))




#scree plot of the first 10 MDS
scree_bac <- scree_plot(bPCoA_bray)
scree_fun <- scree_plot(fPCoA_bray)
fig_scree <- cowplot::plot_grid(scree_bac, scree_fun, align = "hv", nrow=1) #combine

#PLOT
#combine PCoA and scree plot
fig_6.1 <- cowplot::plot_grid(fig_PCoA, fig_scree, align = "hv", nrow=2, rel_heights = c(9,3))

#print
#png("figures/fig6.1.png",12000, 9000, res = 600)
fig_6.1
#dev.off()

```

\vspace{5mm}

### Figure 6.2 \| CAP - genotype:arsenic effect on beta diversity

\vspace{5mm}

```{r CAP genotype arsenic, echo=F, warning=F, eval=T, fig.height=9, fig.width=12}

#BACTERIA
bCAP_bray <- phyloseq::ordinate(bPHYSEQ, method="CAP", ~ genotype * arsenic, distance="bray") #CAP
CAP_bray_var_tbl <- variability_table(bCAP_bray) # extract variation details
phylum.sum <- tapply(phyloseq::taxa_sums(bPHYSEQ), phyloseq::tax_table(bPHYSEQ)[, "labels"], sum, na.rm=T)
topphyla <- names(sort(phylum.sum, T))[1:6]
phy_sub <- prune_taxa((phyloseq::tax_table(bPHYSEQ)[, "labels"] %in% topphyla), bPHYSEQ)
score <- paste("[ ~ genotype * arsenic \n arsenic: ",
               format(bdist_paov$R2[rownames(bdist_paov)=="arsenic"] * 100, digits=2, nsmall=1),
               "% var, P=",format(bdist_paov$`Pr(>F)`[rownames(bdist_paov)=="arsenic"], digits=2),"]",
                "\n genotype: ",
               format(bdist_paov$R2[rownames(bdist_paov)=="genotype"] * 100, digits=2, nsmall=1),
               "% var, P=",format(bdist_paov$`Pr(>F)`[rownames(bdist_paov)=="genotype"], digits=2),"]",
                "\n arsenic:genotype: ",
               format(bdist_paov$R2[rownames(bdist_paov)=="genotype:arsenic"] * 100, digits=2, nsmall=1),
               "% var, P=",format(bdist_paov$`Pr(>F)`[rownames(bdist_paov)=="genotype:arsenic"], digits=2),"]", sep = "")
#plot
p0 <- phyloseq::plot_ordination(phy_sub, bCAP_bray, color="genotype", shape="arsenic", title =  "Bacteria - CAP: Genotype:Arsenic effect", axes=c(1,2))
p0$layers <- p0$layers[-1]
blines <- centroid_lines(phy_sub, bCAP_bray, c("genotype","arsenic"), axes=c(1,2))
p0 <- p0 + geom_segment(aes(xend=blines$Axis.1, yend=blines$Axis.2), x = blines$Axis.1.means, y = blines$Axis.2.means, alpha=0.3, lwd=1)
p0 <- p0 + geom_point(size=3)
#p0 <- p0 + geom_text(aes(label=bPHYSEQ@sam_data$bac_FLD))
p0 <- p0 + scale_color_manual(values=level_cols_genotype)
p0 <- p0 + labs(subtitle = score)
p0 <- p0 + theme_bw()
fig_bac <- p0


#FUNGI
fCAP_bray <- phyloseq::ordinate(fPHYSEQ, method="CAP", ~ genotype * arsenic, distance="bray") #CAP
CAP_bray_var_tbl <- variability_table(fCAP_bray) # extract variation details
phylum.sum <- tapply(phyloseq::taxa_sums(fPHYSEQ), phyloseq::tax_table(fPHYSEQ)[, "labels"], sum, na.rm=T)
topphyla <- names(sort(phylum.sum, T))[1:6]
phy_sub <- prune_taxa((phyloseq::tax_table(fPHYSEQ)[, "labels"] %in% topphyla), fPHYSEQ)
score <- paste("[ ~ genotype * arsenic \n arsenic: ",
               format(fdist_paov$R2[rownames(fdist_paov)=="arsenic"] * 100, digits=2, nsmall=1),
               "% var, P=",format(fdist_paov$`Pr(>F)`[rownames(fdist_paov)=="arsenic"], digits=2),"]",
                "\n genotype: ",
               format(fdist_paov$R2[rownames(fdist_paov)=="genotype"] * 100, digits=2, nsmall=1),
               "% var, P=",format(fdist_paov$`Pr(>F)`[rownames(fdist_paov)=="genotype"], digits=2),"]",
                "\n arsenic:genotype: ",
               format(fdist_paov$R2[rownames(fdist_paov)=="genotype:arsenic"] * 100, digits=2, nsmall=1),
               "% var, P=",format(fdist_paov$`Pr(>F)`[rownames(fdist_paov)=="genotype:arsenic"], digits=2),"]", sep = "")
#plot
p0 <- phyloseq::plot_ordination(phy_sub, fCAP_bray, color="genotype", shape="arsenic", title =  "Fungi - CAP: Genotype:Arsenic effect", axes=c(1,2))
p0$layers <- p0$layers[-1]
flines <- centroid_lines(phy_sub, fCAP_bray, c("genotype","arsenic"))
p0 <- p0 + geom_segment(aes(xend=flines$Axis.1, yend=flines$Axis.2), x = flines$Axis.1.means, y = flines$Axis.2.means, alpha=0.3, lwd=1)
p0 <- p0 + geom_point(size=3)
#p0 <- p0 + geom_text(aes(label=fPHYSEQ@sam_data$fun_FLD))
p0 <- p0 + scale_color_manual(values=level_cols_genotype)
p0 <- p0 + labs(subtitle = score)
p0 <- p0 + theme_bw()
fig_fun <- p0

#combine the two plots (library cowplot)
fig_PCA <- cowplot::plot_grid(fig_bac + theme(legend.position="none"),
               fig_fun + theme(legend.position="none") + labs(colour=""),
               #labels=c("Bacteria","Fungi"),
               align = "hv",
               nrow=1)

#add legend
legend <- cowplot::get_legend(p0 + theme(legend.position="bottom") + labs(fill="Plastic"))
fig_PCA <- cowplot::plot_grid(fig_PCA, legend, nrow=2, rel_heights=c(1, .1))

#save
#png("figs_for_paper/fig6.2_PCA.png",5000, 3200, res = 600)
#fig_PCA
#dev.off()

#scree plot of the first 10 MDS
scree_bac <- scree_plot(bCAP_bray)
scree_fun <- scree_plot(fCAP_bray)
fig_scree <- cowplot::plot_grid(scree_bac, scree_fun, align = "hv", nrow=1) #combine

#PLOT
#combine PCoA and scree plot
fig_6.2 <- cowplot::plot_grid(fig_PCA, fig_scree, align = "hv", nrow=2, rel_heights = c(9,3))
fig_6.2

```

\vspace{5mm}

**Conclusion:** There are differences in the bacterial and fungal communities due to the arsenic treatment, the genotypes and their interactions. We can explain about 4% of bacterial and 3% of fungal variety due to the arsenic:genotype interaction effect.

\pagebreak
